Go Implementation of Techniques for Weighted Language Equivalence

### 2020-10-11 paper/sections/conclusion.tex
finishing paper, added docs

Another idea for the future of this implementation, is the addition
of a \texttt{Semiring} interface type that would permit much more insightful analysis
on weighted automata, not only on the field of real numbers.

>>>>

Another topic worth further investigation is the cause of the 
sudden drop at 1e-15 in the F1 score plots.

An idea for the future of this implementation, is the addition
of a \texttt{Semiring} interface type that would permit much more insightful analysis
on weighted automata, not only on the field of real numbers.

### 2020-10-11 paper/sections/implementation.tex
finishing paper, added docs

To represent the congruence relation $R$, we introduce a \texttt{struct} containing:
\begin{itemize}
    \item A dense matrix \texttt{s} of size $n \times 2k$ containing the set of pairs in the relation
    \item A dense matrix \texttt{u} to represent the generating set $U_R$ for the congruence closure (see definition \ref{def:congclos}).
    \item Two integers representing the number of pairs in the set, and the size of vectors.
    \item A tolerance value to be used in equivalence checks. Best results were obtained with $10^{-14}$.
\end{itemize}

@@

At the end of \texttt{BPR}, we store the basis for the orthogonal complement
of $\llwb$ as an attribute of the automaton. To check if two vectors $v_1 \llwb v_2$,
we check that $R_j^T(v_1 - v_2) = \vec{0}$, with a prefixed tolerance.

@@

\begin{proof}
  Let $V$ be an inner product space over the field $\K$ with an inner product defined as
  $\langle \cdot , \cdot \rangle : V \times V \to \K$. 
  Every linear functional can be 
  represented with a vector. Let $\xi : V \to \K$ be a functional, 
  $\xi \in  W^0$. Because $\xi(w)=0 \quad \forall  w \in W$, 
  if $v$ represents $\xi$ we have that $(v, w)=\xi(w)=0$ for all $w \in W$. 
  We obtain that $\varphi(W^0) \subseteq W^{\perp}$.
  If $v \in W^\perp$  
  then the functional $x \mapsto (v, x)$ cancels over $W$ 
  (by the definition of orthogonality).
\end{proof}

>>>>

To represent the congruence relation $R$, we introduce a \texttt{struct} containing:
\begin{itemize}
    \item A dense matrix \texttt{s} of size $n \times 2k$ containing the set of pairs in the relation
    \item A dense matrix \texttt{u} to represent the generating set $U_R$ for the congruence closure (see definition \ref{def:congclos}).
    \item Two integers representing the number of pairs in the set, and the size of vectors.
    \item A tolerance value to be used in equivalence checks.
\end{itemize}

@@

At the end of \texttt{BPR}, we store $R_j$, the basis for the orthogonal complement
of $\llwb$ as an attribute of the automaton. To check if two vectors $v_1 \llwb v_2$,
we check that $R_j^T(v_1 - v_2) = \vec{0}$, with a given tolerance.

@@

\begin{proof}
  Let $V$ be an inner product space over the field $\K$ with an inner product defined as
  $\langle \cdot , \cdot \rangle : V \times V \to \K$. 
  Every linear functional can be 
  represented with a vector. Let $\xi : V \to \K$ be a functional, 
  $\xi \in  W^0$. Because $\xi(w)=0 \quad \forall  w \in W$, 
  if $v$ represents $\xi$ we have that $\langle v, w \rangle =\xi(w)=0$ for all $w \in W$. 
  We obtain that $\varphi(W^0) \subseteq W^{\perp}$.
  If $v \in W^\perp$  
  then the functional $\xi \mapsto \langle v, x \rangle$ cancels over $W$ 
  (by the definition of orthogonality).
\end{proof}

### 2020-10-11 paper/sections/introduction.tex
finishing paper, added docs

Weighted automata (WAs) are a generalization of non-deterministic automata.
When reading a symbol, a non-deterministic automaton 
can transition in different states simultaneously. Weighted automata
introduce \textit{weights} over transitions, which can for example, 
represent the cost of a transition, 
or in probabilistic systems, the chance of such transition to happen.
WAs can be represented with a set of states, an output function and
a set of transition matrices, indexed over the symbols in the alphabet
the automaton can read. While those automata are typically defined over semirings,
our implementation will focus for simplicity on automata with 
transitions defined over the real number field. 
The current configuration of a finite weighted automaton $W$,
defined on $n$ states, will be represented with a column vector of length
$n$, with values over the semiring or field on which transition weights of the
automaton $W$ are also defined.

>>>>

Weighted automata (WAs) are a generalization of non-deterministic automata.
When reading a symbol, a non-deterministic automaton 
can transition in different states simultaneously. Weighted automata
introduce \textit{weights} over transitions, which can for example, 
represent the cost of a transition, 
or in probabilistic systems, the chance of such transition to happen.
WAs can be represented with a set of states, an output function and
a set of transition matrices, indexed over the symbols in the alphabet
the automaton can read. While those automata are typically defined over semirings,
for simplicity, our implementation will focus only on automata with 
transitions defined over the field of real numbers. 
The current configuration of a finite weighted automaton $W$,
defined on $n$ states, will be represented with a column vector of length
$n$, with values over the semiring or field on which transition weights of the
automaton $W$ are also defined.

### 2020-10-11 paper/sections/notation.tex
finishing paper, added docs

\begin{defn}
  A \textit{linear weighted automaton} (in short, LWA) over the field $\K$ and 
  an alphabet $A$
  is a triple  $L = (V, o, \left\lbrace t_a \right\rbrace_{a \in A})$ 
  where $V$ is a vector space representing the state space, 
  $o: V \to \K$ is a linear map associating to each state its output weight,
  and $t = \left\lbrace t_a \in \K^{V \times V} \right\rbrace_{a \in A}$ is
  the set of transition functions, represented with liner maps 
  that for each input $a \in A$ associate the next state, in this case a vector
  in $V$.
  As seen in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)}$.
\end{defn}

>>>>

\begin{defn}
  A \textit{linear weighted automaton} (in short, LWA) over the field $\K$ and 
  an alphabet $A$
  is a triple  $L = (V, o, \left\lbrace t_a \right\rbrace_{a \in A})$ 
  where $V$ is a vector space representing the state space, and $\dim{(V)} = n$;
  We have that
  $o: V \to \K$ is a linear map associating to each state its output weight,
  and $t = \left\lbrace t_a \in \K^{n \times n} \right\rbrace_{a \in A}$ is
  the set of transition functions, represented with liner maps 
  that for each input $a \in A$ associate the next state, in this case a vector
  in $V$.
  As seen in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)} = n$.
\end{defn}

### 2020-10-11 paper/sections/runtime.tex
finishing paper, added docs

%\subsection{Tolerance and F1 Score}
We the studied of the influence 
of the tolerance value over the correctness of the algorithms results.
Tests took place on randomly generated automata. A basis of  $\llwb$ was 
computed for each automaton with BPR. An arbitrary number of vector pairs were generated 
as an uniformly distributed random linear combination of the vectors in the spanning set of $\llwb$. 
The same quantity of vector pairs was generated randomly with an uniform distribution.
BPREquivalence and HKC procedures were then executed with those pairs in input.

@@

We have introduced \textit{accuracy, precision, recall} and \textit{F1 score} over 
the tests results:

\begin{equation*}
    \begin{aligned}
        \text{Accuracy} = \dfrac{TP + TN}{TP + TN + FP + FN} & \quad & 
        \text{Recall} = \dfrac{TP}{TP + FN} \\ && \\
        \text{Precision} = \dfrac{TP}{TP + FP} & \quad &
        \text{F1} = 2 * \dfrac{\text{Precision} \cdot \text{Recall}}{\text{Precision} + \text{Recall}}  \\ && \\
    \end{aligned}
\end{equation*}

We have then collected 
the F1 score in relation to varying tolerance values.
The plot is shown in Figure \ref{fig:f1}.  
The program was executed on an x86\_64 AMD Ryzen 5 2600X Six Core CPU, running at 3.6GHz with 32GB RAM, 
running Void Linux. Executed sequentially, F1 score tests took 2m26.19s
Parallelized with a worker pool, the tests took 1m13.12s.

% //TODO: condizionamento BPR
TODO: BPR conditioning

@@

TODO empty LLWB chance, memory and time plots, varying automata size parameters

>>>>

%\subsection{Tolerance and F1 Score}
We the studied of the influence 
of the tolerance value over the correctness of the algorithms results.
Tests took place on randomly generated automata. A basis of  $\llwb$ was 
computed for each automaton with BPR. An arbitrary number of vector pairs were generated 
as uniformly distributed random linear combinations of the vectors in the spanning set of $\llwb$. 
The same quantity of vector pairs was generated randomly with an uniform distribution.
BPREquivalence and HKC procedures were then executed with those pairs in input.

@@

We then borrow four concepts from binary classification: \textit{accuracy, precision, recall} 
and \textit{F1 score}:

\begin{equation*}
    \begin{aligned}
        \text{Accuracy} = \dfrac{TP + TN}{TP + TN + FP + FN} & \quad & 
        \text{Recall} = \dfrac{TP}{TP + FN} \\ && \\
        \text{Precision} = \dfrac{TP}{TP + FP} & \quad &
        \text{F1} = 2 \cdot \dfrac{\text{Precision} \cdot \text{Recall}}{\text{Precision} + \text{Recall}}  \\ && \\
    \end{aligned}
\end{equation*}

We have then computed 
the F1 score in relation to varying tolerance values.
The plot is shown in Figures \ref{fig:f1} and \ref{fig:f1-2}.  
The program was executed on an x86\_64 AMD Ryzen 5 2600X Six Core CPU, running at 3.6GHz with 32GB RAM, 
running Void Linux. Executed sequentially, F1 score tests took an average 2m26.19s,
Parallelized by using a worker pool, the tests lasted the times reported in 
Figure \ref{fig:f1} and \ref{fig:f1-2}.
For example, in the first plot of Figure \ref{fig:f1-2}, the test took 82 seconds.
Since we computed BPR for 3000 random automata, over 17 tolerance values, for 3 line plots,
with a chance of finding a non-empty basis for $\llwb$ of approximately $1/3$,
the BPR algorithm was executed an approximate average of
$\frac{3000 \cdot 3 \cdot 17}{82} \approx 1865$ times per second. 
The HKC algorithm, was executed 2000 times on each automaton with a non-empty 
basis of $\llwb$, therefore, it was executed an approximate average of 
$\frac{3000 \cdot 17 \cdot 2000 \cdot 3}{3 \cdot 82} \approx 1243902$ times per second.

% //TODO: condizionamento BPR
%TODO: BPR conditioning

@@

%TODO empty LLWB chance, memory and time plots, varying automata size parameters
